---
title: '575 Final Project: Data Analysis'
author: "Lizzie, Suhaib, Emma"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

# Load packages
```{r load-pkg, message=FALSE}
library(tidyverse)
library(haven)
library(ggplot2)
library(here)
library(jtools)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(modelsummary)
library(knitr)
library(kableExtra)
library(brms)
library(gtsummary)
theme_set(jtools::theme_apa())
```

# Load Data
## Import Data
```{r data-import, message=FALSE}
# Read in the data
df.hatch <- read_sav(here("HATCH 09.29.21.sav"))
```

## Tidy Data
```{r data-tidy, message=FALSE}
# Subset Variables
df.mlm <- df.hatch %>%
  select(c(CoupID,                                # Couple ID
           
           # Couple Level                         ## COUPLE ##
           DelMod, ModeofDeliverySpecific,        # Delivery Method
           GesAgeWk,                              # Gestational age
           Bb.sex,                                # Baby sex

           # Person Level                         ## PERSON ##
           contains("peritot"),                   # BEQ (birth stress)
           contains("pnAge"),                     # Parent age
           contains("Ethn"),                      # Parent ethnicity
           contains("Educ"),                      # Parent level of education
           contains("psam"),                      # Pregnancy-Specific Anxiety Measure
           starts_with("pnBDI"),                  # Prenatal depression
           starts_with("pnPSS"),                  # Prenatal Perceived Stress Scale

           
           # Time Level                           ## TIME ##
           bage3pp.1, bage6pp, bage12pp.1,        # Baby age
           contains("PSI_t"),                     # Parenting Stress
              contains("PSIt"),
              contains("PSI.t"),
           starts_with("BDI"),                    # Depression
           starts_with("PSS"),                    # Perceived Stress Scale
           starts_with("EPDS")                    # Postpartum Depression
           )) %>%
  select(-c(EPDS.12.1, EPDS12.2))

# Rename Variables
colnames(df.mlm) <- c("CoupID",
                      
                      # Couple Level
                      "DelMeth", "DelMeth.Specific",
                      "GestationAge",
                      "BabySex",
                      
                      # Person Level
                      "beq.mom", "beq.dad",
                      "age.mom", "age.dad",
                      "ethnicity.mom", "ethnicity.dad",
                      "education.mom", "education.dad",
                      "pnPSAM.mom", "pnPSAM.dad",
                      "pnBDI.mom", "pnBDI.dad",
                      "pnPSS.mom", "pnPSS.dad",
                      
                      
                      # Time Level
                      "BabyAge.3", "BabyAge.6", "BabyAge.12",
                      "PSI.3.mom", "PSI.12.mom", "PSI.12.dad",
                          "PSI.3.dad", "PSI.6.dad",
                          "PSI.6.mom",
                      "BDI.3.mom", "BDI.3.dad", 
                        "BDI.6.mom", "BDI.6.dad",
                        "BDI.12.mom", "BDI.12.dad",
                      "PSS.3.mom", "PSS.3.dad", 
                        "PSS.6.mom", "PSS.6.dad",
                        "PSS.12.mom", "PSS.12.dad",
                      "EPDS.3.mom", "EPDS.3.dad", 
                        "EPDS.6.mom", "EPDS.6.dad",
                        "EPDS.12.mom", "EPDS.12.dad"
                      )

# Reorder PSI Variables
df.mlm <- df.mlm %>%
  relocate(c("PSI.3.mom", "PSI.3.dad",
             "PSI.6.mom", "PSI.6.dad",
             "PSI.12.mom", "PSI.12.dad"),
           .after = BabyAge.12) 
```

## Wide to Long 
```{r wide-to-long}
# Person-Level and Time-Level Variables
df.long1 <- df.mlm %>%
  # Parent-level variables
  pivot_longer(
    cols = beq.mom:pnPSS.dad,
    names_to = c(".value", "parent"),
    names_pattern = "(beq|age|ethnicity|education|pnPSAM|pnBDI|pnPSS).(mom|dad)",
    names_transform = list(parent = as.factor)
  ) %>%
  # Assign participant IDs
  mutate(PersonID = seq(1:200)) %>%
  # Time-level variables
  pivot_longer(
    cols = BabyAge.3:BabyAge.12,
    names_to = c(".value", "time"),
    names_pattern = "(BabyAge).(3|6|12)",
    names_transform = list(time = as.integer)
  ) %>%
  # Remove time-varying variables
  select(-starts_with(c("PSI", "BDI", "PSS", "EPDS")))

# PSI Variable
df.long2 <- df.mlm %>%
  pivot_longer(
    cols = PSI.3.mom:EPDS.12.dad,
    names_to = c(".value", "time", "parent"),
    names_pattern = "(PSI|BDI|PSS|EPDS).(3|6|12).(mom|dad)",
    names_transform = list(time = as.integer, parent = as.factor)
  ) %>%
  # Select relevant variables
  select(c(CoupID, time, parent, PSI, BDI, PSS, EPDS))


# Combine Data Frames
df.long <- left_join(df.long1, df.long2, by = c("CoupID", "parent", "time")) %>%
  # Move identifying variable
  relocate(c(PersonID, parent, time), .after = CoupID) %>%
  # Remove funky formatting
  mutate(beq = na_if(beq, -97.450))
```

## Data Processing
```{r, message=FALSE}
#scaling and decomposing of variables
df_scaled <- df.long %>% 
  mutate(log_BDI = log(BDI+1)) %>%
  mutate(DelMeth = as.factor(DelMeth)) %>%
  mutate(log_EPDS = log(EPDS+1)) %>%
  group_by(CoupID) %>% mutate(beq_cm = mean(beq)) %>% ungroup() %>% mutate(beq_cmc = beq - beq_cm, beq_st = scale(beq), beq_stc = scale(beq_cmc)) 
  
```

## Save Data
```{r}
write_sav(df_scaled, here("../df_scaled.sav"))
```

# Plots
## Density Plots
```{r plot-density, message=FALSE}
ggplot(data = df.long, aes(x = BDI)) +
  geom_density() +
  facet_grid(parent~time) +
  labs(x = "Depression (BDI)", y = "")

ggplot(data = df_scaled, aes(x = log_BDI)) +
  geom_density() +
  facet_grid(parent~time) +
  labs(x = "Log-Transformed Depression (BDI)", y = "")

ggplot(data = df.long, aes(x = EPDS)) +
  geom_density() +
  facet_grid(parent~time) +
  labs(x = "PP Depression (EPDS)", y = "")

ggplot(data = df.long, aes(x = beq)) + 
  geom_density() + 
  facet_wrap(~parent) +
  labs(x = "Birthing Stress (BEQ)", y = "")
```

## Bar Plot
```{r plot-bar}
ggplot(data = distinct(df.long, CoupID, .keep_all = T), 
       aes(x = as.factor(DelMeth), 
           fill = as.factor(DelMeth.Specific))) + 
  geom_bar(position = "stack", color = "black") + 
  scale_x_discrete(labels = c("Vaginal", "C-Section", "Missing")) + 
  scale_fill_discrete(labels = c("Natural Vaginal", "Medicated Vaginal", "Unplanned C-Section", "Planned C-Section", "Missing")) +
  labs(x = "Delivery Method", y = "")
```

## Spaghetti Plot
```{r}
ggplot(df.long, aes(x = time, y = BDI, color = factor(CoupID))) + 
  geom_jitter(size = 0.5, alpha = 0.2, width = 0.1, height = 0.1) + 
  geom_line(aes(group = PersonID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  stat_summary(fun = "mean", col = "black", size = 1.5, geom = "line") + 
  facet_wrap(~parent) + 
   # Suppress legend
  guides(color = "none", fill = "none") +
  # Labels
  labs(title = "Change in Depression Levels", 
       x = "", y = "BDI") + 
  # Remove Label
  scale_x_continuous(breaks = c(3,6, 12), 
                     labels = c("3PP", "6PP", "12PP"))
```

## Scatterplots 
```{r}
ggplot(df_scaled, aes(x = beq_cmc, y = BDI, color = factor(parent))) +
  geom_jitter() +
  geom_smooth(method = "lm", se = F) + 
  facet_grid(~time)
```

## Box Plots
```{r}
ggstatsplot::grouped_ggbetweenstats(data = df_scaled, 
               x = time,
               y = BDI,
               grouping.var = parent,
               outlier.tagging = T,
               outlier.label = CoupID,
               centrality.point.args = list(size = 4),
               plot.type = "box",
               results.subtitle = F,
               pairwise.comparisons = F,
               annotation.args = list(title = ""),
               ylab = "", xlab = ""
                 )
               
```

# Models

## Random Slope Tests
```{r message=FALSE}
m01 <- lmer(log_BDI ~ 1 + (time|CoupID/PersonID), data = df_scaled, control = lmerControl(optimizer = "nmkbw")) 
m02 <- lmer(log_BDI ~ 1 + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = lmerControl(optimizer = "nmkbw")) 
ranova(m01) #coup -> significant, person nonsignificant -> did not converge so test again without random slope on Person but random intercept
ranova(m02) #signficiant and converges!
```

## Basic Models
```{r}
# Raw BDI
m1 <- glmmTMB(BDI ~ beq_cmc + time*DelMeth*beq_cm + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled)
summary(m1)

# Log-Transformed BDI, include both random slopes for convergence! Remove the 3way interaction!
m2 <- glmmTMB(log_BDI ~ beq_cmc + time*beq_cm + beq_cm*DelMeth + time*DelMeth + (time|CoupID/PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m2)

# Log-Transformed BDI without time interaction, does not converge with random slope, so take out!
m3 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + (1|CoupID/PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m3)

# EPDS
m3.5 <- glmmTMB(log_EPDS ~ beq_cmc + time*DelMeth*beq_cm + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m3.5)
```
Model 1 looks promising:

1. Delivery method predicts BDI: Delivery method 1 is associated with a significant increase in BDI
2. couple-level beq predicts BDI: Couple level beq (what exactly is it measuring?) is associated with a significant increase in BDI
3. BDI is not impacted by time-since-birth
4. There is an interaction of DelMeth and beq, with DelMeth1 reducing the influence of beq on BDI, and this interaction is stable over time!
5. The impact of delivery method does NOT change over time
6. The impact of couple-level beq does NOT change over time 
7. Individual-level beq does NOT predict BDI

### Covariate Models
```{r}
# Gender
m4 <- glmmTMB(BDI ~ beq_cmc + time*DelMeth*beq_cm + parent + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled)
summary(m4)

# I think it is better to keep all data in the dataset and account for gender, instead of filtering out!

# # Gender and Prenatal Depression -> does not converge
# m5 <- glmmTMB(log_BDI ~ beq_cmc + time*DelMeth*beq_cm + parent + pnBDI + (1|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
# summary(m5)
# 
# # Gender, pnBDI, parent-level demographics, does not converge
# m6 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + ethnicity + age + education + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="Nelder-Mead")))
# summary(m6)
# 
# # Gender, pnBDI, baby demographics, does not converge
# m7 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + GestationAge + BabySex + (time|PersonID) + (1|CoupID), data = df_scaled)
# summary(m7)

#Models do not converge -> use bayesian models below instead!!
```

## Interaction Plot
```{r}
interactions::interact_plot(m4, 
              pred = "beq_cm",
              modx = "DelMeth",
              plot.points = TRUE, 
              point.size = 0.5, 
              point.alpha = 0.2, 
              facet.modx = FALSE)
```
### Bayesian

```{r}
# Same syntax; just change `lmer()` to `brm()`
m7_robust <- brm(
  log_BDI ~ beq_cmc + time*DelMeth*beq_cm + parent + pnBDI + GestationAge + BabySex + (time|CoupID) + (1|CoupID:PersonID),
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m7_robust)
```
Significant:
  *DelMeth1                 
  *pnBDI                    
  *DelMeth1:beq_cm


```{r}
# Same syntax; just change `lmer()` to `brm()`
m6_robust <- brm(
  log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + ethnicity + age + education +(time|CoupID) + (1|CoupID:PersonID),
    # Assume random effects follow a t distribution
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m6_robust)
```
Significant:
  *time               
  *DelMeth
  *beq_cm 
  *pnBDI
  *DelMeth1:beq_cm


## Model Equation:
Nomenclature:  

* k: Couples
* j: persons
* i: time intervals

### Level 1: Time

$$\text{logBDI}_{ijk} = \beta_{0jk} + \beta_{1jk} \text{time}_{ijk} + e_{ijk}$$

### Level 2: Person

$$
  \begin{aligned}
    \beta_{0jk} & = \gamma_{00k} + \gamma_{01k}\text{BEQ_cmc}_{jk} + \gamma_{02k}\text{gender}_{jk} + \gamma_{03k}\text{pnBDI}_{jk}  + u_{0jk}\\
    \beta_{1jk} & = \gamma_{10k} + u_{1jk}
  \end{aligned}
$$
### Level 3: Couple 
$$
  \begin{aligned}
    \gamma_{00k} & = \gamma_{000} + \gamma_{001}\text{DelMeth}_k + \gamma_{002}\text{BEQ_cm}_k + \gamma_{003}\text{DelMeth}_k\times\text{BEQ_cm}_k  \\
    \gamma_{01k} & = \gamma_{010} \\
    \gamma_{02k} & = \gamma_{020} \\
    \gamma_{03k} & = \gamma_{030} \\
    \gamma_{10k} & = \gamma_{100} 
  \end{aligned}
$$
### Combined
$$
  \begin{aligned}
    \text{logBDI}_{ijk} & = \gamma_{000} + \gamma_{001}\text{DelMeth}_k + \gamma_{002}\text{BEQ_cm}_k + \gamma_{003}\text{DelMeth}_k\times\text{BEQ_cm}_k \\
    & + \gamma_{010}\text{BEQ_cmc}_{jk} + \gamma_{020}\text{gender}_{jk} + \gamma_{030}\text{pnBDI}_{jk}\\
    &+ \gamma_{100}\text{time}_{ijk} \\
    &+  u_{1jk}\text{time}_{ijk} + u_{0jk} + e_{ijk}
  \end{aligned}
$$

# Tables
## Predictor Names -> needs to be adjusted to the chosen models!
```{r include=FALSE}
names <- c(
  '(Intercept)' = "Intercept",
  'time' = "Time Since Birth",
  'beq_cmc' = "Person-Level Birth Stress",
  'beq_cm' = "Couple-Centered Birth Stress",
  'DelMeth1' = "Delivery Method (C-Section)",
  'DelMeth1:beq_cm' = "C-Section X Couple-Center Birth Stress",
  'parentmom' = "Gender (Female)",
  'pnBDI' = "Prenatal Depression",
  'time:beq_cm' = "Time Since Birth X Couple-Center Birth Stress",
  'time:DelMeth1' = "Time Since Birth X C-Section",
  'time:DelMeth1:beq_cm' = "Time Since Birth X C-Section X Couple-Center Birth Stress"
)

order <- c(
  '(Intercept)',
  'time',
  'beq_cmc',
  'beq_cm',
  'DelMeth1',
  'DelMeth1:beq_cm',
  'time:DelMeth1',
  'time:beq_cm',
  'time:DelMeth1:beq_cm',
  'parentmom',
  'pnBDI'
)
```

## msummary
```{r eval=FALSE, message=FALSE, include=FALSE}
msummary(list("3-Way Interaction Model" = m2,
              "2-Way Interaction Model" = m3,
              "Covariate Model"= m5),
         group = group + term ~ model,
         fmt = 2,
         title = "Table 1",
         coef_rename = names,
         estimate = "{estimate} [{conf.low}, {conf.high}] {stars}",
         statistic = NULL,
         stars = TRUE,
         output = "kableExtra",
         add_rows = new_rows)

new_rows <- data.frame(
  term = c("Fixed effects", "Random effects", "Goodness of Fit"),
  m2 = c("", "", ""),
  m3 = c("", "", ""),
  m5 = c("", "", ""),
  m0 = c("", "", "")
)

attr(new_rows, "position") <- c(1, 13, 19)

# Step 1. Create a map for the predictors and the terms in the model.
# Need to use '\\( \\)' to show math.
# Rename and reorder the rows. Need to use '\\( \\)' to
# show math. If this does not work for you, don't worry about it.
cm <- c(names,
        "PersonID_sd__(Intercept)" = "\\(\\tau_0\\)",
        "sd__time" = "\\(\\tau_1\\) (time)",
        "sd__Observation" = "\\(\\sigma\\)")
# Step 2. Create a list of models to have one column for coef, 
# one for SE, one for CI, and one for p
models <- list(
  "Estimate" = m2,
  "95% CI" = m2,
)
# Step 3. Add rows to say fixed and random effects
# (Need same number of columns as the table)
new_rows <- data.frame(
  term = c("Fixed effects", "Random effects"),
  est = c("", ""),
  ci = c("", "")
)
# Specify which rows to add
attr(new_rows, "position") <- c(1, 11)
# Step 4. Render the table
msummary(models,
         estimate = c("estimate", 
                      "[{conf.low}, {conf.high}]"),
         stars = T,
         statistic = NULL, 
         coef_map = cm,
         add_rows = new_rows)
```

## sjPlot
```{r}
sjPlot::tab_model(m2, m3, m5,
                  dv.labels = c("3-Way Interaction Model", "2-Way Interaction Model", "Covariate Model"),
                  show.aic = T,
                  show.loglik = T,
                  string.ci = "95% CI",
                  pred.labels = names)
```

