---
title: '575 Final Project: Data Analysis'
author: "Lizzie, Suhaib, Emma"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

# Load packages
```{r load-pkg, message=FALSE}
library(tidyverse)
library(haven)
library(ggplot2)
library(here)
library(jtools)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(modelsummary)
library(knitr)
library(kableExtra)
library(brms)
library(gtsummary)
theme_set(jtools::theme_apa())
uscred <- "#990000"
uscgold <- "#ffcc00"
```

# Load Data
## Import Data
```{r data-import, message=FALSE}
df.hatch <- read_sav(here("HATCH 09.29.21.sav")) # original .sav file
df_scaled <- read_sav(here("df_scaled.sav")) # tidied data from `mlm_datacleaning.rmd`
```

# Testing
## ICC
```{r}
m0 <- lmer(beq ~ 1 + (1|CoupID/PersonID), data = df_scaled)
variance_components <- as.data.frame(VarCorr(m0))
between_var <- variance_components$vcov[1]
within_var <- variance_components$vcov[2]
(icc <- between_var / (between_var + within_var))
```

## Random Slope Tests
```{r rnd-slp}
m01 <- lmer(log_BDI ~ 1 + (time|CoupID/PersonID), data = df_scaled, control = lmerControl(optimizer = "nmkbw")) 
m02 <- lmer(log_BDI ~ 1 + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = lmerControl(optimizer = "nmkbw")) 
ranova(m01) #coup -> significant, person nonsignificant -> did not converge so test again without random slope on Person but random intercept
ranova(m02) #significant and converges!
```

# Models 
## Basic Models
```{r model-basics}
# 3 Way Interaction
m1 <- glmmTMB(log_BDI ~ beq_cmc + time*beq_cm*DelMeth + (time|CoupID) + (1|CoupID:PersonID), 
              data = df_scaled, 
              control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m1)

# 2 Way Interaction
m2 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + (time|CoupID) + (1|CoupID:PersonID), 
              data = df_scaled, 
              control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m2)
```


1. Delivery method predicts BDI: Delivery method 1 is associated with a significant increase in BDI
2. Couple-level beq predicts BDI: Couple level beq (what exactly is it measuring?) is associated with a significant increase in BDI
3. BDI *may* not be impacted by time-since-birth
4. There is an interaction of DelMeth and beq, with DelMeth1 reducing the influence of beq on BDI, and this interaction is stable over time
5. The impact of delivery method does NOT change over time
6. The impact of couple-level beq does NOT change over time 
7. Individual-level beq does NOT predict BDI

## Covariate Models
```{r}
# Relevant Covariates
m3 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + (time|CoupID) + (1|CoupID:PersonID),
              data = df_scaled, 
              control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m3)

# Demographics Covariates
m4 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + GestationAge + BabySex + education + ethnicity + (time|CoupID) + (1|CoupID:PersonID), 
              data = df_scaled, 
              control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m4)

# Models do not converge -> use bayesian models below instead!!
```

## Bayesian Models 
### Run Models
```{r model-bayesian-run, eval=FALSE, include=FALSE}
# Model 1 as bayesian
m1_robust <- brm(
  log_BDI ~ beq_cmc + time*beq_cm*DelMeth + (time|CoupID:PersonID), # includes random slopes for both; m1 only included random slopes for couple
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)

save(m1_robust, file = here("m1_robust.Rdata"))

# Model 2 as bayesian
m2_robust <- brm(
  log_BDI ~ beq_cmc + time + DelMeth*beq_cm + (time|CoupID/PersonID),
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)

save(m2_robust, file = here("m2_robust.Rdata"))

# M3 as Bayesian
m3_robust <- brm(
  log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + (time|CoupID/PersonID),
    # Assume random effects follow a t distribution
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)

save(m3_robust, file = here("m3_robust.Rdata"))

# M4 as Bayesian
m4_robust <- brm(
  log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + GestationAge + BabySex + education + ethnicity + (time|CoupID/PersonID),
    # Assume random effects follow a t distribution
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)

save(m4_robust, file = here("m4_robust.Rdata"))
```

### Interpret Models
#### Three-Way Interaction
```{r m1-bayes}
load(here("m1_robust.Rdata"))
summary(m1_robust)
```
Significant:
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *beq_cm: increases BDI
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm
  
#### Two-Way Interaction
```{r m2-bayes}
load(here("m2_robust.Rdata"))
summary(m2_robust)
```
Significant:
  *time: reduction of BDI over time
  *beq_cm: increases BDI
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm

#### Two-Way Interaction with Relevant Covariates
```{r m3-bayes}
load(here("m3_robust.Rdata"))
summary(m3_robust)
```
Significant:
  *time: reduction of BDI
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *beq_cm: increases BDI
  *pnBDI: increases BDI
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm

#### Two-Way Interaction with Demographic Covariates
```{r m4-bayes}
load(here("m4_robust.Rdata"))
summary(m4_robust)
```
Significant:
  *time: reduces BDI
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *beq_cm: increases BDI
  *pnBDI: increases BDI
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm


# Model Equations for Model 3:

**Nomenclature:**  

* k: Couples
* j: persons
* i: time intervals

**Level 1: Time**

$$\text{logBDI}_{ijk} = \beta_{0jk} + \beta_{1jk} \text{time}_{ijk} + e_{ijk}$$

**Level 2: Person**

$$
  \begin{aligned}
    \beta_{0jk} & = \gamma_{00k} + \gamma_{01k}\text{BEQ_cmc}_{jk} + \gamma_{02k}\text{gender}_{jk} + \gamma_{03k}\text{pnBDI}_{jk}  + u_{0jk}\\
    \beta_{1jk} & = \gamma_{10k} + u_{1jk}
  \end{aligned}
$$
**Level 3: Couple** 

$$
  \begin{aligned}
    \gamma_{00k} & = \gamma_{000} + \gamma_{001}\text{DelMeth}_k + \gamma_{002}\text{BEQ_cm}_k + \gamma_{003}\text{DelMeth}_k\times\text{BEQ_cm}_k + v_{00k} \\
    \gamma_{01k} & = \gamma_{010}  + v_{01k}\\
    \gamma_{02k} & = \gamma_{020}  + v_{02k}\\
    \gamma_{03k} & = \gamma_{030}  + v_{03k}\\
    \gamma_{10k} & = \gamma_{100}  + v_{10k}
  \end{aligned}
$$


# Tables
## Demographics
```{r}
t1 <- df_scaled %>%
  group_by(PersonID) %>% mutate(bdi_mean = mean(BDI, na.rm = T)) %>% ungroup() %>%
  distinct(PersonID, .keep_all = T) %>%
  mutate(ethnicity = factor(x = ethnicity, levels = c(1:7), labels = c("White", "Black", "Hispanic/LatinX", "American Indian/Alaska Native", "AAPI", "Other", "Declined"))) %>% 
  mutate(education = factor(x = education, levels = c(1:6), labels = c("High School/GED", "Some College", "Associate's Degree", "Bachelor's Degree", "Master's Degree", "Professional/Doctoral Degree"))) %>%
  select(age, beq, pnBDI, bdi_mean, ethnicity, education, parent) %>%
  tbl_summary(statistic = all_continuous() ~ "{mean} ({sd})",
              missing = "no", 
              by = parent,
              label = list(age ~ "Parent Age", 
                           beq ~ "BEQ",
                           pnBDI ~ "Prenatal BDI",
                           bdi_mean ~ "Postpartum BDI",
                           ethnicity ~ "Ethnicity",
                           education ~ "Education")) %>%
  add_overall() %>%
  italicize_levels() %>%
  modify_header(list(stat_0 ~ "**All Parents**", 
                     stat_1 ~ "**Fathers**",
                     stat_2 ~ "**Mothers**"))

t2 <- df_scaled %>%
  distinct(CoupID, .keep_all = T) %>%
  mutate(BabySex = factor(x = BabySex, levels = c(0, 1), labels = c("Female", "Male"))) %>%
  mutate(DelMeth = factor(x = DelMeth, levels = c(1,2), labels = c("Vaginal", "C-Section"))) %>%
  select(GestationAge, BabySex, DelMeth) %>%
  tbl_summary(statistic = all_continuous() ~ "{mean} ({sd})", 
              missing = "no") %>%
  italicize_levels()

tbl_stack(list(t1, t2)) %>%
  as_kable_extra(caption = "Table 1. Demographics") %>% 
  kable_classic(full_width = F) %>%
  pack_rows("Level 2", 1, 19) %>%
  pack_rows("Level 3", 20, 26)

## Missing Values
map(distinct(df_scaled, CoupID, .keep_all = T), ~sum(is.na(.)))
map(filter(df_scaled, time == 12), ~sum(is.na(.)))

```

# Plots
## Density Plots
```{r plot-density, message=FALSE}

df_scaled <- df_scaled %>%
  mutate(time2 = factor(time, levels = c(3, 6, 12), labels = c("3 PP", "6 PP", "12 PP")))

# BDI Density
f1 <- ggplot(data = df_scaled, aes(x = BDI, fill=as.factor(parent))) +
  geom_density(alpha=0.6, show.legend = FALSE) +
  facet_grid(.~time2, scales = "free", space='free') +
  scale_fill_manual(values=c(uscred, uscgold)) +
  labs(x = "Depression (BDI)", y = "",
       title = "Figure 1", subtitle = "Distribution of BDI")

f2 <- ggplot(data = df_scaled, aes(x = log_BDI, fill=as.factor(parent))) +
  geom_density(alpha=0.6) +
  facet_grid(.~time2) +
  scale_fill_manual(labels = c("Dad", "Mom"), values=c(uscred, uscgold)) + 
  labs(x = "Log-Transformed Depression (BDI)", y = "",
       title = "Figure 2", subtitle = "Distribution of Log-Transformed BDI")

gridExtra::grid.arrange(f1, f2, ncol = 2)

# BEQ
ggplot(data = df_scaled, aes(x = beq, fill = as.factor(parent))) + 
  geom_density(alpha = 0.6) +
  scale_fill_manual(labels = c("Dad", "Mom"), values=c(uscred, uscgold)) + 
  labs(x = "Birthing Stress (BEQ)", y = "",
       title = "Figure 3", subtitle = "Distribution of BEQ") 

```


## Bar Plot
```{r plot-bar}
# Delivery Method Without Missing Data
ggplot(data = distinct(na.omit(df_scaled), CoupID, .keep_all = T), 
       aes(x = as.factor(DelMeth), 
           fill = as.factor(DelMeth.Specific))) + 
  geom_bar(position = "stack", color = "black") + 
  scale_x_discrete(labels = c("Vaginal", "C-Section")) + 
  scale_fill_discrete(labels = c("Natural Vaginal", "Medicated Vaginal", "Unplanned C-Section", "Planned C-Section")) +
  labs(x = "Delivery Method", y = "",
       title = "Figure 4", subtitle = "Types of Delivery")
```

## Spaghetti Plot
```{r}
df_scaled <- df_scaled %>%
  mutate(parent2 = factor(parent, levels = c(1:2), labels = c("Fathers", "Mothers")))

ggplot(df_scaled, aes(x = time, y = BDI, color = factor(CoupID))) + 
  geom_jitter(size = 0.5, alpha = 0.2, width = 0.1, height = 0.1) + 
  geom_line(aes(group = PersonID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  stat_summary(fun = "mean", col = "black", size = 1.5, geom = "line") + 
  facet_wrap(~parent2) + 
   # Suppress legend
  guides(color = "none", fill = "none") +
  # Labels
  labs(title = "Figure 5", subtitle = "Change in Depression Levels", 
       x = "", y = "BDI") + 
  # Remove Label
  scale_x_continuous(breaks = c(3,6, 12), 
                     labels = c("3PP", "6PP", "12PP")) 

```

## Scatterplots 
```{r}
ggplot(df_scaled, aes(x = beq_cmc, y = BDI, color = factor(parent))) +
  geom_jitter(alpha = 0.6) +
  geom_smooth(method = "lm", se = F) + 
  facet_grid(~time2) +   
  scale_color_manual(labels = c("Dad", "Mom"), values=c(uscred, uscgold)) + 
  labs(title = "Figure 6", subtitle = "Depression ~ Person-Level Birth Stress", 
       x = "Person-Level Birth Stress", y = "BDI")

```

## Interaction Plot
```{r}
df_csv <- df_scaled %>%
  mutate(DelMeth = factor(DelMeth, levels = c(1:2), labels = c("0", "1")))

m_interact <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + (time|CoupID) + (1|CoupID:PersonID), 
              data = df_csv, 
              control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))

interactions::interact_plot(m_interact, 
              pred = "beq_cm",
              modx = "DelMeth",
              plot.points = TRUE, 
              point.size = 0.5, 
              point.alpha = 0.2, 
              facet.modx = FALSE,
              x.label = "Couple-Level Subjective Birth Stress",
              y.label = "Log-Tranformed BDI",
              legend.main = "Delivery Method",
              modx.labels = c("Vaginal", "C-Section"),
              main.title = "Figure 7 
Interaction Plot")

m_future <- glmmTMB(log_BDI ~ beq_cm + time + DelMeth*beq_cmc*parent + (time|CoupID) + (1|CoupID:PersonID), 
              data = df_csv, 
              control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))

interactions::interact_plot(m_future, 
              pred = "beq_cmc",
              modx = "DelMeth",
              mod2 = "parent",
              plot.points = TRUE, 
              point.size = 0.5, 
              point.alpha = 0.2, 
              facet.modx = FALSE, 
              x.label = "Individual-Level Subjective Birth Stress",
              y.label = "Log-Tranformed BDI",
              legend.main = "Delivery Method",
              modx.labels = c("Vaginal", "C-Section"),
              mod2.labels = c("Father", "Mother"),
              main.title = "Figure 8 
Future Directions")
```
