---
title: '575 Final Project: Data Analysis'
author: "Lizzie, Suhaib, Emma"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

# Load packages
```{r load-pkg, message=FALSE}
library(tidyverse)
library(haven)
library(ggplot2)
library(here)
library(jtools)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(modelsummary)
library(knitr)
library(kableExtra)
library(brms)
# library(gtsummary)
theme_set(jtools::theme_apa())
# theme_set(theme_default())
```

# Load Data
## Import Data
```{r data-import, message=FALSE}
df.hatch <- read_sav(here("../HATCH 09.29.21.sav")) # original .sav file
df_scaled <- read_sav(here("../df_scaled.sav")) # tidied data from `mlm_datacleaning.rmd`
```

# Plots
## Density Plots
```{r plot-density, message=FALSE}
# BDI Density
f1 <- ggplot(data = df_scaled, aes(x = BDI, fill=as.factor(parent))) +
  geom_density(alpha=0.6, show.legend = FALSE) +
  facet_grid(.~time, scales = "free", space='free') +
  labs(x = "Depression (BDI)", y = "") + scale_fill_manual(values=c("#F8766D", "#00BA38")) #change colors here

f2 <- ggplot(data = df_scaled, aes(x = log_BDI, fill=as.factor(parent))) +
  geom_density(alpha=0.6) +
  facet_grid(.~time) +
  labs(x = "Log-Transformed Depression (BDI)", y = "") + scale_fill_manual(values=c("#F8766D", "#00BA38")) #change colors here

# BEQ
ggplot(data = df_scaled, aes(x = beq, fill = as.factor(parent))) + 
  geom_density(alpha = 0.6) +
  labs(x = "Birthing Stress (BEQ)", y = "",
       title = "Figure 3", subtitle = "Distribution of BEQ") + scale_fill_manual(values=c("#F8766D", "#00BA38")) #change colors here

```
```{r fig.height=4, fig.width=16}
# Combine BDI Density Plots
gridExtra::grid.arrange(f1, f2, ncol = 2)
```


## Bar Plot
```{r plot-bar}
# Delivery Method Without Missing Data
ggplot(data = distinct(na.omit(df_scaled), CoupID, .keep_all = T), 
       aes(x = as.factor(DelMeth), 
           fill = as.factor(DelMeth.Specific))) + 
  geom_bar(position = "stack", color = "black") + 
  scale_x_discrete(labels = c("Vaginal", "C-Section")) + 
  scale_fill_discrete(labels = c("Natural Vaginal", "Medicated Vaginal", "Unplanned C-Section", "Planned C-Section")) +
  labs(x = "Delivery Method", y = "",
       title = "Figure 4", subtitle = "Types of Delivery")

# Delivery Method With Missing Data
# ggplot(data = distinct(df.long, CoupID, .keep_all = T), 
#       aes(x = as.factor(DelMeth), 
#           fill = as.factor(DelMeth.Specific))) + 
#  geom_bar(position = "stack", color = "black") + 
#  scale_x_discrete(labels = c("Vaginal", "C-Section", "Missing")) + 
#  scale_fill_discrete(labels = c("Natural Vaginal", "Medicated Vaginal", "Unplanned C-Section", "Planned C-Section", "Missing")) +
#  labs(x = "Delivery Method", y = "")
```

## Spaghetti Plot
```{r}
ggplot(df_scaled, aes(x = time, y = BDI, color = factor(CoupID))) + 
  geom_jitter(size = 0.5, alpha = 0.2, width = 0.1, height = 0.1) + 
  geom_line(aes(group = PersonID)) +  # add lines to connect the data for each person
  # add a mean trajectory
  stat_summary(fun = "mean", col = "black", size = 1.5, geom = "line") + 
  facet_wrap(~parent) + 
   # Suppress legend
  guides(color = "none", fill = "none") +
  # Labels
  labs(title = "Figure 5", subtitle = "Change in Depression Levels", 
       x = "", y = "BDI") + 
  # Remove Label
  scale_x_continuous(breaks = c(3,6, 12), 
                     labels = c("3PP", "6PP", "12PP"))
```

## Scatterplots -- NEEDS FORMATTING
```{r}
ggplot(df_scaled, aes(x = beq_cmc, y = BDI, color = factor(parent))) +
  geom_jitter() +
  geom_smooth(method = "lm", se = F) + 
  facet_grid(~time) +
```

## Box Plots -- NEEDS FORMATTING
```{r}
ggstatsplot::grouped_ggbetweenstats(data = df_scaled, 
               x = time,
               y = BDI,
               grouping.var = parent,
               outlier.tagging = T,
               outlier.label = CoupID,
               centrality.point.args = list(size = 4),
               plot.type = "box",
               results.subtitle = F,
               pairwise.comparisons = F,
               annotation.args = list(title = ""),
               ylab = "", xlab = ""
                 )
               
```

## Interaction Plot
```{r}
#why is it here? it needs to run after the respective model!
interactions::interact_plot(m5, 
              pred = "beq_cm",
              modx = "DelMeth",
              plot.points = TRUE, 
              point.size = 0.5, 
              point.alpha = 0.2, 
              facet.modx = FALSE)
```

# Models

## Random Slope Tests
```{r message=FALSE}
m01 <- lmer(log_BDI ~ 1 + (time|CoupID/PersonID), data = df_scaled, control = lmerControl(optimizer = "nmkbw")) 
m02 <- lmer(log_BDI ~ 1 + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = lmerControl(optimizer = "nmkbw")) 
ranova(m01) #coup -> significant, person nonsignificant -> did not converge so test again without random slope on Person but random intercept
ranova(m02) #significant and converges!
```

## Basic Models
```{r}
# Raw BDI, remove time interactions?
m1 <- glmmTMB(BDI ~ beq_cmc + time*DelMeth*beq_cm + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled)
summary(m1)

# Log-Transformed BDI, include both random slopes for convergence! Remove the 3way interaction!
m2 <- glmmTMB(log_BDI ~ beq_cmc + time*beq_cm + beq_cm*DelMeth + time*DelMeth + (time|CoupID/PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m2)

# Log-Transformed BDI without time interaction, does not converge with random slope, so take out!
m3 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + (1|CoupID/PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
summary(m3)
```

Model 1 and 2 look promising:

1. Delivery method predicts BDI: Delivery method 1 is associated with a significant increase in BDI
2. Couple-level beq predicts BDI: Couple level beq (what exactly is it measuring?) is associated with a significant increase in BDI
3. BDI is not impacted by time-since-birth
4. There is an interaction of DelMeth and beq, with DelMeth1 reducing the influence of beq on BDI, and this interaction is stable over time!
5. The impact of delivery method does NOT change over time
6. The impact of couple-level beq does NOT change over time 
7. Individual-level beq does NOT predict BDI

### Covariate Models
```{r}
# Gender, remove time interactions?
m4 <- glmmTMB(BDI ~ beq_cmc + time*DelMeth*beq_cm + parent + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled)
summary(m4)

# I think it is better to keep all data in the dataset and account for gender, instead of filtering out!

# # Gender and Prenatal Depression -> does not converge
# m5 <- glmmTMB(log_BDI ~ beq_cmc + time*DelMeth*beq_cm + parent + pnBDI + (1|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="BFGS")))
# summary(m5)
# 
# # Gender, pnBDI, parent-level demographics, does not converge
# m6 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + ethnicity + age + education + (time|CoupID) + (1|CoupID:PersonID), data = df_scaled, control = glmmTMBControl(optimizer = optim, optArgs = list(method="Nelder-Mead")))
# summary(m6)
# 
# # Gender, pnBDI, baby demographics, does not converge
# m7 <- glmmTMB(log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + GestationAge + BabySex + (time|PersonID) + (1|CoupID), data = df_scaled)
# summary(m7)

#Models do not converge -> use bayesian models below instead!!
```

## Interaction Plot
```{r}
interactions::interact_plot(m4, 
              pred = "beq_cm",
              modx = "DelMeth",
              plot.points = TRUE, 
              point.size = 0.5, 
              point.alpha = 0.2, 
              facet.modx = FALSE)
```

# Bayesian

```{r}
# Model 4 as bayesian, remove time interactions? below is a version without!
m4_robust <- brm(
  log_BDI ~ beq_cmc + time*DelMeth + beq_cm*DelMeth + time*beq_cm + parent + (time|CoupID/PersonID),
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m4_robust)
```
Significant:
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *beq_cm: increases BDI
  *parent: Mom increases BDI compared to dad
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm
  
  
```{r}
# m4 including baby demographics, no time interactions
m5_robust <- brm(
  log_BDI ~ beq_cmc + time + beq_cm*DelMeth + parent + pnBDI + GestationAge + BabySex + (time|CoupID/PersonID),
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m5_robust)
```

Significant:
  *time: reduction of BDI over time
  *beq_cm: increases BDI
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *pnBDI: increases BDI (slightly?)
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm
  
```{r}
# m4 including parent demographics, no time interaction
m6_robust <- brm(
  log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + pnBDI + ethnicity + age + education + (time|CoupID/PersonID),
    # Assume random effects follow a t distribution
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m6_robust)
```

Significant:
  *time: reduction of BDI
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *beq_cm: increases BDI
  *pnBDI: increases BDI
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm

```{r}
# m4 including all covariates, include time interactions? Version without below!
m7_robust <- brm(
  log_BDI ~ beq_cmc + time*DelMeth + time*beq_cm + DelMeth*beq_cm + parent + GestationAge + pnBDI + ethnicity + age + education + BabySex + (time|CoupID/PersonID),
    # Assume random effects follow a t distribution
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m7_robust)
```

Significant:
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *pnBDI: increases BDI
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm

```{r}
# m4 including all covariates, no time interactions
m8_robust <- brm(
  log_BDI ~ beq_cmc + time + DelMeth*beq_cm + parent + GestationAge + pnBDI + ethnicity + age + education + BabySex + (time|CoupID/PersonID),
    # Assume random effects follow a t distribution
  # Assume the errors follow a t distribution
  family = student(),
  data = df_scaled,
  chains = 2,
  iter = 4000,
  control = list(adapt_delta = .95)
)
summary(m8_robust)
```

Significant:
  *time: reduces BDI
  *DelMeth1: DelMeth1 increases BDI compared to DelMeth0
  *beq_cm: increases BDI
  *pnBDI: increases BDI
  *DelMeth1:beq_cm: DelMeth1 reduces the effect of beq_cm
  
## Model Equation:
Nomenclature:  

* k: Couples
* j: persons
* i: time intervals

### Level 1: Time

$$\text{logBDI}_{ijk} = \beta_{0jk} + \beta_{1jk} \text{time}_{ijk} + e_{ijk}$$

### Level 2: Person

$$
  \begin{aligned}
    \beta_{0jk} & = \gamma_{00k} + \gamma_{01k}\text{BEQ_cmc}_{jk} + \gamma_{02k}\text{gender}_{jk} + \gamma_{03k}\text{pnBDI}_{jk}  + u_{0jk}\\
    \beta_{1jk} & = \gamma_{10k} + u_{1jk}
  \end{aligned}
$$
### Level 3: Couple 
$$
  \begin{aligned}
    \gamma_{00k} & = \gamma_{000} + \gamma_{001}\text{DelMeth}_k + \gamma_{002}\text{BEQ_cm}_k + \gamma_{003}\text{DelMeth}_k\times\text{BEQ_cm}_k  \\
    \gamma_{01k} & = \gamma_{010} \\
    \gamma_{02k} & = \gamma_{020} \\
    \gamma_{03k} & = \gamma_{030} \\
    \gamma_{10k} & = \gamma_{100} 
  \end{aligned}
$$
### Combined
$$
  \begin{aligned}
    \text{logBDI}_{ijk} & = \gamma_{000} + \gamma_{001}\text{DelMeth}_k + \gamma_{002}\text{BEQ_cm}_k + \gamma_{003}\text{DelMeth}_k\times\text{BEQ_cm}_k \\
    & + \gamma_{010}\text{BEQ_cmc}_{jk} + \gamma_{020}\text{gender}_{jk} + \gamma_{030}\text{pnBDI}_{jk}\\
    &+ \gamma_{100}\text{time}_{ijk} \\
    &+  u_{1jk}\text{time}_{ijk} + u_{0jk} + e_{ijk}
  \end{aligned}
$$

# Tables
## Predictor Names -> needs to be adjusted to the chosen models!
```{r include=FALSE}
names <- c(
  '(Intercept)' = "Intercept",
  'time' = "Time Since Birth",
  'beq_cmc' = "Person-Level Birth Stress",
  'beq_cm' = "Couple-Centered Birth Stress",
  'DelMeth1' = "Delivery Method (C-Section)",
  'DelMeth1:beq_cm' = "C-Section X Couple-Center Birth Stress",
  'parentmom' = "Gender (Female)",
  'pnBDI' = "Prenatal Depression",
  'time:beq_cm' = "Time Since Birth X Couple-Center Birth Stress",
  'time:DelMeth1' = "Time Since Birth X C-Section",
  'time:DelMeth1:beq_cm' = "Time Since Birth X C-Section X Couple-Center Birth Stress"
)

order <- c(
  '(Intercept)',
  'time',
  'beq_cmc',
  'beq_cm',
  'DelMeth1',
  'DelMeth1:beq_cm',
  'time:DelMeth1',
  'time:beq_cm',
  'time:DelMeth1:beq_cm',
  'parentmom',
  'pnBDI'
)
```

## msummary
```{r eval=FALSE, message=FALSE, include=FALSE}
msummary(list("3-Way Interaction Model" = m2,
              "2-Way Interaction Model" = m3,
              "Covariate Model"= m5),
         group = group + term ~ model,
         fmt = 2,
         title = "Table 1",
         coef_rename = names,
         estimate = "{estimate} [{conf.low}, {conf.high}] {stars}",
         statistic = NULL,
         stars = TRUE,
         output = "kableExtra",
         add_rows = new_rows)

new_rows <- data.frame(
  term = c("Fixed effects", "Random effects", "Goodness of Fit"),
  m2 = c("", "", ""),
  m3 = c("", "", ""),
  m5 = c("", "", ""),
  m0 = c("", "", "")
)

attr(new_rows, "position") <- c(1, 13, 19)

# Step 1. Create a map for the predictors and the terms in the model.
# Need to use '\\( \\)' to show math.
# Rename and reorder the rows. Need to use '\\( \\)' to
# show math. If this does not work for you, don't worry about it.
cm <- c(names,
        "PersonID_sd__(Intercept)" = "\\(\\tau_0\\)",
        "sd__time" = "\\(\\tau_1\\) (time)",
        "sd__Observation" = "\\(\\sigma\\)")
# Step 2. Create a list of models to have one column for coef, 
# one for SE, one for CI, and one for p
models <- list(
  "Estimate" = m2,
  "95% CI" = m2,
)
# Step 3. Add rows to say fixed and random effects
# (Need same number of columns as the table)
new_rows <- data.frame(
  term = c("Fixed effects", "Random effects"),
  est = c("", ""),
  ci = c("", "")
)
# Specify which rows to add
attr(new_rows, "position") <- c(1, 11)
# Step 4. Render the table
msummary(models,
         estimate = c("estimate", 
                      "[{conf.low}, {conf.high}]"),
         stars = T,
         statistic = NULL, 
         coef_map = cm,
         add_rows = new_rows)
```

## sjPlot
```{r}
sjPlot::tab_model(m6_robust, m7_robust)
```

## Demographics
```{r}
t1 <- df_scaled %>%
  group_by(PersonID) %>% mutate(bdi_mean = mean(BDI, na.rm = T)) %>% ungroup() %>%
  distinct(PersonID, .keep_all = T) %>%
  
  mutate(ethnicity = factor(x = ethnicity, levels = c(1:7), labels = c("White", "Black", "Hispanic/LatinX", "American Indian/Alaska Native", "AAPI", "Other", "Declined"))) %>% 
  mutate(education = factor(x = education, levels = c(1:6), labels = c("High School/GED", "Some College", "Associate's Degree", "Bachelor's Degree", "Master's Degree", "Professional/Doctoral Degree"))) %>%
  
  select(age, beq, pnBDI, bdi_mean, ethnicity, education, parent) %>%
  tbl_summary(statistic = all_continuous() ~ "{mean} ({sd})",
              missing = "no", 
              by = parent,
              label = list(age ~ "Parent Age", 
                           beq ~ "BEQ",
                           pnBDI ~ "Prenatal BDI",
                           bdi_mean ~ "Postpartum BDI",
                           ethnicity ~ "Ethnicity",
                           education ~ "Education")) %>%
  add_overall() %>%
  italicize_levels() %>%
  modify_header(list(stat_0 ~ "**All Parents**", 
                     stat_1 ~ "**Fathers**",
                     stat_2 ~ "**Mothers**"))

t2 <- df_scaled %>%
  distinct(CoupID, .keep_all = T) %>%
  mutate(BabySex = factor(x = BabySex, levels = c(0, 1), labels = c("Female", "Male"))) %>%
  mutate(DelMeth = factor(x = DelMeth, levels = c(0, 1), labels = c("Vaginal", "C-Section"))) %>%
  select(GestationAge, BabySex, DelMeth) %>%
  tbl_summary(statistic = all_continuous() ~ "{mean} ({sd})", 
              missing = "no") %>%
  italicize_levels()

tbl_stack(list(t1, t2)) %>%
  as_kable_extra(caption = "Table 1") %>% 
  kable_classic(full_width = F) %>%
  pack_rows("Level 2", 1, 19) %>%
  pack_rows("Level 3", 20, 26)

```

