---
title: '575 Final Project: Data Analysis'
author: "Lizzie, Suhaib, Emma"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(tinytex.verbose = TRUE)
```

# Load packages
```{r load-pkg, message=FALSE}
library(tidyverse)
library(haven)
library(ggplot2)
library(here)
library(jtools)
library(lme4)
library(lmerTest)
library(glmmTMB)
library(modelsummary)
library(knitr)
library(kableExtra)
theme_set(jtools::theme_apa())
```

# Load Data
## Import Data
```{r data-import, message=FALSE}
# Read in the data
df.hatch <- read_sav(here("HATCH 09.29.21.sav"))
```

## Tidy Data
```{r data-tidy, message=FALSE}
c# Subset Variables
df.mlm <- df.hatch %>%
  select(c(CoupID,                                # Couple ID
           
           # Couple Level                         ## COUPLE ##
           DelMod, ModeofDeliverySpecific,        # Delivery Method
           GesAgeWk,                              # Gestational age
           Bb.sex,                                # Baby sex
           
           # Person Level                         ## PERSON ##
           contains("peritot"),                   # BEQ (birth stress)
           contains("pnAge"),                     # Parent age
           contains("Ethn"),                      # Parent ethnicity
           contains("Educ"),                      # Parent level of education
           contains("psam"),                      # Pregnancy-Specific Anxiety Measure
           starts_with("pnBDI"),                  # Prenatal depression
           starts_with("pnPSS"),                  # Prenatal Perceived Stress Scale
           
           # Time Level                           ## TIME ##
           bage3pp.1, bage6pp, bage12pp.1,        # Baby age
           contains("PSI_t"),                     # Parenting Stress
              contains("PSIt"),
              contains("PSI.t"),
           starts_with("BDI"),                    # Depression
           starts_with("PSS")                     # Perceived Stress Scale
           ))   

# Rename Variables
colnames(df.mlm) <- c("CoupID",
                      
                      # Couple Level
                      "DelMeth", "DelMeth.Specific",
                      "GestationAge",
                      "BabySex",
                      
                      # Person Level
                      "beq.mom", "beq.dad",
                      "age.mom", "age.dad",
                      "ethnicity.mom", "ethnicity.dad",
                      "education.mom", "education.dad",
                      "pnPSAM.mom", "pnPSAM.dad",
                      "pnBDI.mom", "pnBDI.dad",
                      "pnPSS.mom", "pnPSS.dad",
                      
                      
                      # Time Level
                      "BabyAge.3", "BabyAge.6", "BabyAge.12",
                      "PSI.3.mom", "PSI.12.mom", "PSI.12.dad",
                          "PSI.3.dad", "PSI.6.dad",
                          "PSI.6.mom",
                      "BDI.3.mom", "BDI.3.dad", 
                        "BDI.6.mom", "BDI.6.dad",
                        "BDI.12.mom", "BDI.12.dad",
                      "PSS.3.mom", "PSS.3.dad", 
                        "PSS.6.mom", "PSS.6.dad",
                        "PSS.12.mom", "PSS.12.dad"
                      )

# Reorder PSI Variables
df.mlm <- df.mlm %>%
  relocate(c("PSI.3.mom", "PSI.3.dad",
             "PSI.6.mom", "PSI.6.dad",
             "PSI.12.mom", "PSI.12.dad"),
           .after = BabyAge.12)
```

## Wide to Long 
```{r wide-to-long}
# Person-Level and Time-Level Variables
df.long1 <- df.mlm %>%
  # Parent-level variables
  pivot_longer(
    cols = beq.mom:pnPSS.dad,
    names_to = c(".value", "parent"),
    names_pattern = "(beq|age|ethnicity|education|pnPSAM|pnBDI|pnPSS).(mom|dad)",
    names_transform = list(parent = as.factor)
  ) %>%
  # Assign participant IDs
  mutate(PersonID = seq(1:200)) %>%
  # Time-level variables
  pivot_longer(
    cols = BabyAge.3:BabyAge.12,
    names_to = c(".value", "time"),
    names_pattern = "(BabyAge).(3|6|12)",
    names_transform = list(time = as.integer)
  ) %>%
  # Remove PSI variables
  select(-starts_with(c("PSI", "BDI", "PSS")))

# PSI Variable
df.long2 <- df.mlm %>%
  pivot_longer(
    cols = PSI.3.mom:PSS.12.dad,
    names_to = c(".value", "time", "parent"),
    names_pattern = "(PSI|BDI|PSS).(3|6|12).(mom|dad)",
    names_transform = list(time = as.integer, parent = as.factor)
  ) %>%
  # Remove variables in `df.long1` that aren't identifying variables
  select(c(CoupID, time, parent, PSI, BDI, PSS))


# Combine Data Frames
df.long <- left_join(df.long1, df.long2, by = c("CoupID", "parent", "time")) %>%
  # Move identifying variable
  relocate(c(PersonID, parent, time), .after = CoupID) %>%
  # Remove funky formatting
  mutate(beq = na_if(beq, -97.450))
```


# Density Plots
```{r plot-density, message=FALSE}
ggplot(data = df.long, aes(x = PSI)) +
  geom_density() +
  facet_grid(parent~time) +
  labs(x = "Parenting Stress (PSI)", y = "")

ggplot(data = df.long, aes(x = beq)) + 
  geom_density() + 
  facet_wrap(~parent) +
  labs(x = "Birthing Stress (BEQ)", y = "")

ggplot(data = na.omit(distinct(df.long, CoupID, .keep_all = T)), 
       aes(x = as.factor(DelMeth), 
           fill = as.factor(DelMeth.Specific))) + 
  geom_bar(position = "stack", color = "black") + 
  scale_x_discrete(labels = c("Vaginal", "C-Section")) + 
  scale_fill_discrete(labels = c("Natural Vaginal", "Medicated Vaginal", "Unplanned C-Section", "Planned C-Section")) +
  labs(x = "Delivery Method", y = "")
```


## Data Processing

```{r, message=FALSE}
#scaling and decomposing of variables
df_scaled <- df.long %>% mutate(log_psi = scale(log(PSI))) %>%
  group_by(CoupID) %>% mutate(beq_cm = mean(beq)) %>% ungroup() %>% mutate(beq_cmc = beq - beq_cm)
```

## Models

### Model Equation:
Nomenclature:
j: Couples
i: persons
t: time intervals

\begin{align}
\label{eq:m2}
psi_{tij} = \gamma_{000} + \gamma_{100} \cdot TIME_{tij} + \gamma_{010}\cdot beq^{cmc}_{ij} + u_{0i0} + \gamma_{001}\cdot beq^{cm}_{j} + \gamma_{002}\cdot DelMeth_j + \gamma_{003}\cdot DelMeth\cdot beq^{cm}_j + u_{00j} + e_{tij}
\end{align}

### Modelling
```{r message=FALSE}
m0 <- lmer(PSI ~ 1 + (time|CoupID), data=df_scaled)
ranova(m0) #not significant!
#random slope insignificant for Couples and does not converge for Persons
```

The random slope on the couple level is not significant, so it will not be included in the model.
Note, that the model did not converge for person-level random slopes, which is why they have been excluded for now, however, they are planned to be included in future versions.

```{r}
m1 <- glmmTMB(PSI ~ time + beq_cmc  + DelMeth*beq_cm + (1|PersonID) + (1|CoupID), data = df_scaled)
summary(m1) #robust se estimation (faster than brms)

# m2 <- glmmTMB(PSI ~ time + beq_cmc  + DelMeth*beq_cm + age + parent + (1|PersonID) + (time|CoupID), data = df_scaled)
# summary(m2) #robust se estimation (faster than brms)
```

```{r echo=F}
#function to create table caption
tn = local({
  i = 0
  function(x) {
    # i <<- i + 1
    i = "2.1" #manual numbering to adjust to documents style
    paste('Table ', i, ': ', x, sep = '')
    # The : before Table tells pandoc to wrap your caption in <caption></caption>
  }
})
knit_hooks$set(tab.cap = function(before, options, envir) {
  if(!before)
    tn(options$tab.cap)
})
default_output_hook = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (is.null(options$tab.cap) == F)  
    x
  else
    default_output_hook(x,options)
})
```

```{r table1, tab.cap="Table shows the intercept, level-1 fixed effects (time), level-2 fixed effects (beq_cmc), level-3 fixed effects (beq_cm, delivery method & interaction), random intercept (variance of intercept) and measures of fit (e.g., RMSE).", message=FALSE}
modelsummary(m1, group = group + term ~ model, title = "Model 1 Coefficient Summary") %>% kable_styling()
```

## Plots


```{r figure1, fig.cap="Scatter plot with regression line and 95% confidence interval of parental stress over birth experience (cluster mean centered). Plot shows no significant relation between both variables.", warning=F, message=F}

effect_plot(m1, pred = beq_cmc,interval = TRUE, plot.points = TRUE)

# ggplot(data = df_scaled, aes(x = beq, y=PSI, group=CoupID))+
#   geom_point()+
#   geom_smooth(method = "lm", se = F, aes(colour = CoupID))+
#   xlab("Birth experience")+ylab("Parental Stress")+
#   theme(legend.position = "none")
```


## Interpretation

We investigated the model described in equation \ref{eq:m2}. The model output in table \ref{tab:table1} shows the fixed effects as follows:
For time (level 1) as 0.25 (SE=0.198), for birth experience (cmc, level 2) as -0.04 (SE=2.5), for birth experience (cm, level 3) as 1.8 (SE=3.5), for delivery method (level 3) as 28.5 (SE=25.2) and the interaction of delivery method and birth experience as -4.3 (SE=6.5). As such, none of the main effects were significant. The random intercept for the person-level was 10.9 and on the couple level 13. Figure @\ref{fig:figure1} emphasizes this by showing the connection between birth experience (cmc) and parental stress. The plot shows that there is no significant relation.

These preliminary findings/analysis do not support our hypothesis that birth experience and delivery method relate to post partum parental stress. We will look further into our modelling to be sure that the findings are not caused by our model choices. It could be the case that previous research found relations between post partum depression and birth experience or delivery methods due to unsuitable models (e.g. OLS isntead of MLMs).


 

